A DPG method for steady viscous compressible flow

### 2013-07-30 docs/Jesse/Dissertation/problemRange.tex

In 2D, the classical form of the Navier-Stokes equations involve the fluid density $\rho$, velocity in the $x$ and $y$ directions $u$ and $v$ (or $u_1$ and $u_2$), respectively, temperature $T$, energy per unit mass $e$, and stress and heat flux vectors $\boldsymbol \sigma_i$ and $\vec{q}$. The equations are as follows:
\begin{itemize}
\item{\textbf{Mass conservation}}
\begin{align*}
\pd{\rho}{t} + \div \vecttwo{\rho u }{\rho v} &= 0
\end{align*}

\item{\textbf{Momentum conservation}}
\begin{align*}
\pd{\rho u_1}{t} + \div \left(\vecttwo{\rho u^2+p }{\rho u v} - \boldsymbol \sigma_{1}\right) &=0\\
\pd{\rho u_2}{t} + \div \left(\vecttwo{\rho u v}{\rho v^2+p } - \boldsymbol \sigma_{2}\right) &=0
\end{align*}

\item{\textbf{Energy conservation}}
\begin{align*}
\pd{\rho e}{t} + \div \left(\vecttwo{((\rho e)+p)u}{((\rho e)+p)v} - \boldsymbol \sigma_1 \cdot \boldsymbol u- \boldsymbol \sigma_2 \cdot \boldsymbol u + \vec{q}\right) &=0
\end{align*}
\end{itemize}

We assume our fluid satisfies standard stress laws for $\boldsymbol \sigma$ and $\boldsymbol q$ as well. For viscous stresses $\boldsymbol \sigma$, we assume a Newtonian fluid
\begin{align*}
\sigma_{ij} &= \mu(u_{i,j}+u_{j,i}) + \lambda u_{k,k}\delta_{ij}.
\end{align*}
The coefficients $\lambda$ and $\mu$ are the viscosity and bulk viscosity, respectively. The bulk viscosity is often set implicitly through $2\mu + 3\lambda = 0$, known as Stokes' hypothesis. However, since the effect of bulk viscosity can become important for compressible flows, we treat both coefficients separately. In general, $\mu$ and $\lambda$ are functions of temperature, obeying the power law
\[
\mu = \left(\frac{T}{T_0}\right)^\beta,
\]
where $T_0$ is a reference temperature. We choose $\beta = 2/3$ in this case.

@@

The linearized Navier-Stokes equations are the result of small perturbation assumptions applied to the full Navier-Stokes equations. Under such assumptions, the flow in a domain consists only of slight variations (to a given background flow) that are small compared to the magnitude of the free stream velocity.  Mathematically speaking, the linearized Navier-Stokes equations are the results of the linearization of the full equations with respect to a specific background flow.  The linearized NS equations are frequently referred to as an ``acoustic problem'', as they are used to study sound propagation.

>>>>

In 2D, the classical form of the Navier-Stokes equations involve the fluid density $\rho$, velocity in the $x$ and $y$ directions $u_1$ and $u_2$, respectively, temperature $T$, energy per unit mass $e$, and stress and heat flux vectors $\boldsymbol \sigma_i$ and $\vec{q}$. The equations are as follows:
\begin{itemize}
\item{\textbf{Mass conservation}}
\begin{align*}
\pd{\rho}{t} + \div \vecttwo{\rho u_1 }{\rho u_2} &= 0
\end{align*}

\item{\textbf{Momentum conservation}}
\begin{align*}
\pd{\rho u_1}{t} + \div \left(\vecttwo{\rho u_1^2+p }{\rho u_1 u_2} - \boldsymbol \sigma_{1}\right) &=0\\
\pd{\rho u_2}{t} + \div \left(\vecttwo{\rho u_1 u_2}{\rho u_2^2+p } - \boldsymbol \sigma_{2}\right) &=0
\end{align*}

\item{\textbf{Energy conservation}}
\begin{align*}
\pd{\rho e}{t} + \div \left(\vecttwo{((\rho e)+p)u_1}{((\rho e)+p)u_2} - \boldsymbol \sigma_1 \cdot \boldsymbol u- \boldsymbol \sigma_2 \cdot \boldsymbol u + \vec{q}\right) &=0
\end{align*}
\end{itemize}

We assume our fluid satisfies standard stress laws for $\boldsymbol \sigma$ and $\boldsymbol q$ as well. For viscous stresses $\boldsymbol \sigma$, we assume a Newtonian fluid
\begin{align*}
\sigma_{ij} &= \mu(u_{i,j}+u_{j,i}) + \lambda u_{k,k}\delta_{ij}.
\end{align*}
The coefficients $\lambda$ and $\mu$ are the viscosity and bulk viscosity, respectively. The bulk viscosity is often set implicitly through $2\mu + 3\lambda = 0$, known as Stokes' hypothesis. However, since the effect of bulk viscosity can become important for compressible flows, we treat both coefficients separately. In general, $\mu$ and $\lambda$ are functions of temperature.  One method of modeling temperature dependence is through the power law
\[
\mu = \left(\frac{T}{T_0}\right)^\beta,
\]
where $T_0$ is a reference temperature. We choose $\beta = 2/3$ in this case.

@@

The linearized Navier-Stokes equations are the result of small perturbation assumptions applied to the full Navier-Stokes equations. Under such assumptions, the flow in a domain consists only of slight variations (to a given background flow) that are small compared to the magnitude of the free stream velocity.  Mathematically speaking, the linearized Navier-Stokes equations are the results of the linearization of the full equations with respect to a specific background flow.  %The linearized NS equations are frequently referred to as an ``acoustic problem'', as they are used to study sound propagation.

### 2013-07-30 docs/Jesse/NavierStokes paper/BurgersNotes.tex

\section{A viscous Burgers equation}

@@

Recall that we did not linearize in the flux variables $\widehat{u}$ and $\widehat{f}_n$, so we can directly apply the nonlinear boundary conditions to our variational formulation.  
Additionally, since Burgers' equation does not have any physical constraints, we can employ a direct Newton iteration to solve the nonlinear equation.  The adaptivity algorithm is identical to the greedy algorithm described previously, except that the linear solve is replaced by a nonlinear solve.  The results of an adaptive simulation are shown in Figure~\ref{fig:BurgersShock} for $\epsilon = 1e-4$.  
\begin{figure}[!h]
\centering
\includegraphics[scale=.4]{figs/burgers.png}
\caption{Shock solution for Burgers' equation with $\epsilon = 10^{-4}$.} 
\label{fig:BurgersShock}
\end{figure}

For the solution of Burgers' equation, adaptivity began on a cubic $4\times 4$ mesh.  9 iterations of adaptive mesh refinement were performed, and the resulting mesh and zoomed solution are displayed in Figure~\ref{fig:BurgersShockZoom}, demonstrating a fully resolved smooth transition of the solution into a smeared shock.

\begin{figure}[!h]
\centering
\subfigure{\includegraphics[scale=.275]{figs/burgersZoom.png}}
\subfigure{\includegraphics[scale=.275]{figs/burgersMesh.png}}
\caption{Adaptive mesh after 9 refinements, and zoom view at point (.5,.5) with shock formation and $1e-3$ width line for reference.}
\label{fig:BurgersShockZoom}
\end{figure}

>>>>

\subsection{A viscous Burgers equation}

@@

%\begin{figure}
%\centering
%\caption{Shock solution for Burgers' equation with $\epsilon = 10^{-4}$.} 
%\label{fig:BurgersShock}
%\end{figure}

\begin{figure}
\centering
\subfigure[Solution $u$]{\includegraphics[height=2in]{burgers.png}}
\subfigure[Adaptive mesh]{\includegraphics[height=2in]{burgersMesh.png}}
\subfigure[Zoomed solution at the shock]{\includegraphics[height=2in]{burgersZoom.png}}
\caption{Solution and adaptive mesh after 9 refinements, and zoom view at point (.5,.5) with shock formation and $1e-3$ width line for reference.}
\label{fig:BurgersShock}
\end{figure}

Recall that we did not linearize in the flux variables $\widehat{u}$ and $\widehat{f}_n$, so we can directly apply the nonlinear boundary conditions to our variational formulation.  
Additionally, since Burgers' equation does not have any physical constraints, we can employ a direct Newton iteration to solve the nonlinear equation.  The adaptivity algorithm is identical to the greedy algorithm described previously, except that the linear solve is replaced by a nonlinear solve.  The results of an adaptive simulation are shown in Figure~\ref{fig:BurgersShock} for $\epsilon = 1e-4$.

For the solution of Burgers' equation, adaptivity began on a cubic $4\times 4$ mesh.  9 iterations of adaptive mesh refinement were performed, and the resulting mesh and zoomed solution are also displayed in Figure~\ref{fig:BurgersShock}, demonstrating a fully resolved smooth transition of the solution into a smeared shock.
